From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-09 Pi+/paper/equivalence.tex
Futher work

We present the types and 1-combinators of $\PiPlusLang$ and $\PiHatLang$ in~\Cref*{fig:piplus,fig:pihat} respectively,
eliding the 2-combinators for brevity. We enforce that there is a unique 2-combinator between compatible 1-combinators,
by relating them with a truncation. These can be found in the appendix and in the accompanying Agda code
(~\cref{app:leveltwo}).

@@

First, we show how to translate $\PiLang$ programs to $\PiPlusLang$, which is the additive fragment of $\PiLang$. The
syntax is given in~\Cref{fig:piplus}.

@@

Using this rig structure, we translate $\PiLang$ to $\PiPlusLang$, constructing a rig equivalence from $\PiCat$ to
$\PiPlusCat$.

@@

As described, $\PiHatLang$ doesn't have a tensor product, but we can build it simply by adding up natural numbers, and,
we need to verify that this indeed equips $\PiHatCat$ with a symmetric monoidal structure. Since each object is a
natural number generated by $\zero$ and $\suc$, this makes $\PiHatCat$ similar to a \emph{PROP}, that is, a products and
permutations category.

@@

Using this symmetric monoidal structure, we translate from $\PiPlusLang$ to $\PiHatLang$.

@@

To go back from $\PiHatLang$ to $\PiPlusLang$, we turn a natural number into a $\PiPlusLang$ type, using right-biased
addition, that is, the natural number $n$ gets mapped to the type $\onet + (\onet + (\onet + \ldots + \zerot))$. The
non-trivial part is coherently generating a $\PiPlusLang$ combinator from $\PiHatLang$ combinators, using adjacent swaps
to generate the braiding.

@@

\begin{definition}[${\gdenot{\blank}}$]
  \begin{gather*}
    \begin{aligned}
      \gdenot{\blank}_{0} &: U \to \UFin \\
      \gdenot{\blank}_{0} &\defeq \evalh_{0} \comp \evalp_{0} \comp \evalt_{0}
    \end{aligned}
    \qquad
    \begin{aligned}
      \gdenot{\blank}_{0} &: X \iso Y \to \gdenot{X}_{0} \id_{\UFin} \gdenot{Y}_{0} \\
      \gdenot{\blank}_{1} &\defeq \evalh_{1} \comp \evalp_{1} \comp \evalt_{1}
    \end{aligned}
  \end{gather*}
\end{definition}

>>>>

We present the types and 1-combinators of $\PiPlusLang$ and $\PiHatLang$ in~\Cref*{fig:piplus,fig:pihat} respectively,
eliding the 2-combinators for brevity. We enforce that there is a unique 2-combinator between compatible 1-combinators,
by relating them with a truncation. These can be found in~\cref{app:leveltwo} and in the accompanying Agda code.

@@

First, we show how to translate $\PiLang$ programs to $\PiPlusLang$, which is the additive fragment of $\PiLang$. The
syntax for 1-combinators is given in~\Cref{fig:piplus}.

@@

\noindent Using this rig structure, we translate $\PiLang$ to $\PiPlusLang$, constructing a rig equivalence from
$\PiCat$ to $\PiPlusCat$.

@@

As described, $\PiHatLang$ doesn't have a tensor product, but we can build it simply by adding up natural numbers, and,
we need to verify that this indeed equips $\PiHatCat$ with a symmetric monoidal structure.~\footnote{Since each object
is a natural number, this makes $\PiHatCat$ a \emph{PROP}, that is, a products and permutations category.}

To produce a braiding $n + m \isoh m + n$ from adjacent tranpositions, we recursively traverse the left subexpression,
swapping each element using adjacent transpositions along the elements on the right, placing it in the right position.
The computational content of this translation can be visualised using tree transformations, for the recursive case
in~\cref{fig:plusplusswap}. The challenging part is showing that these moves are coherent with respect to 2-combinators.

\begin{figure}
   \[
      \Tree [ [ {\tiny 1} {\tiny n} ] {\tiny m} ] ~\xrightarrow{}~
      \Tree [ {\tiny 1} [ {\tiny n} {\tiny m} ] ] ~\xrightarrow{}~
      \Tree [ [ {\tiny n} {\tiny m} ] {\tiny 1} ] ~\xrightarrow{}~
      \Tree [ [ {\tiny m} {\tiny n} ] {\tiny 1} ] ~\xrightarrow{}~
      \Tree [ {\tiny m} [ {\tiny n} {\tiny 1} ] ] ~\xrightarrow{}~
      \Tree [ {\tiny m} [ {\tiny 1} {\tiny n} ] ] ~
    \]
  \label{fig:plusplusswap}
  \caption{Braiding from transpositions, recursive case}
\end{figure}

@@

\noindent Using this symmetric monoidal structure, we translate from $\PiPlusLang$ to $\PiHatLang$.

@@

To go back from $\PiHatLang$ to $\PiPlusLang$, we turn a natural number into a $\PiPlusLang$ type, using right-biased
addition, that is, the natural number $n$ gets mapped to the type $\onet + (\onet + (\onet + \ldots + \zerot))$. Since
the types are already right-biased, an adjacent transposition in $\PiHatLang$ is easily encoded by using the braiding in
$\PiPlusLang$.

\begin{figure}
  \[
     \Tree [ {\tiny I} [ {\tiny I} {\tiny n} ] ] ~\xrightarrow{\assoclp}~
     \Tree [ [ {\tiny I} {\tiny I} ] {\tiny n} ] ~\xrightarrow[\phantom{xx}\swapp]{\idc\phantom{xx}}~
     \Tree [ [ {\tiny I} {\tiny I} ] {\tiny n} ] ~\xrightarrow{\assocrp}~
     \Tree [ {\tiny I} [ {\tiny I} {\tiny n} ] ] ~
   \]
 \label{fig:plusplusswap}
 \caption{Transpositions from braiding}
\end{figure}

@@

\begin{definition}[${\gdenot{\blank}}$]
  \begin{gather*}
    \begin{aligned}
      \gdenot{\blank}_{0} & : U \to \UFin                                       \\
      \gdenot{\blank}_{0} & \defeq \evalh_{0} \comp \evalp_{0} \comp \evalt_{0}
    \end{aligned}
    \qquad
    \begin{aligned}
      \gdenot{\blank}_{0} & : X \iso Y \to \gdenot{X}_{0} \id_{\UFin} \gdenot{Y}_{0} \\
      \gdenot{\blank}_{1} & \defeq \evalh_{1} \comp \evalp_{1} \comp \evalt_{1}
    \end{aligned}
  \end{gather*}
\end{definition}

### 2021-07-09 Pi+/paper/examples.tex
Revise sections 6 and 7 and copy-editing
Futher work

\section{Reversible Circuits in Qiskit}
\label{sec:qiskit}

Classical reversible boolean circuits are at the core of most quantum algorithms and
hence are supported by popular platforms for quantum computing such as IBM
Qiskit. Specifically, the Qiskit framework provides the following universal set
of gates for reversible computing: \textsf{not} (boolean negation, called
\verb|x|), \textsf{cnot} (conditional negation of the second input if the first
is true; called \verb|cx|), and \textsf{toffoli} (conditional negation of the
third input if both the first two inputs are true; called \verb|ccx|)
gates. Additionally, Qiskit allows implicit re-shuffling of bits by allowing
each operation to specify the indices of its input bits.

>>>>

\section{Reversible Circuits in Qiskit}
\label{sec:qiskit}
\label{sec:examples}

Classical reversible boolean circuits are at the core of most quantum algorithms and hence are supported by popular
platforms for quantum computing such as IBM Qiskit~\cite{aleksandrowiczQiskitOpensourceFramework2019}. Specifically, the
Qiskit framework provides the following universal set of gates for reversible computing: \textsf{not} (boolean negation,
called \verb|x|), \textsf{cnot} (conditional negation of the second input if the first is true; called \verb|cx|), and
\textsf{toffoli} (conditional negation of the third input if both the first two inputs are true; called \verb|ccx|)
gates. Additionally, Qiskit allows implicit re-shuffling of bits by allowing each operation to specify the indices of
its input bits.

### 2021-07-09 Pi+/paper/finite.tex
Fixes
Futher work

We will show that it is possible to encode all $\PiLang$ combinators using adjacent transpositions
(in~\cref{sec:equivalence}). Group presentations are built by adding equations to a \emph{free group}, which we define
first.

@@

Before, the only way to decide the equality of two elements in a group was to evaluate and check them on the nose, but
in a group presentation, this is reduced to deciding whether one word -- a representative of the equivalence class of
the group's elements, can be reduced to another word, using the group's relations. However, these equations are not
directed, so it is not always possible to construct a well-behaved rewriting system. In general, the word problem for
groups is known to be undecidable.

\subsubsection{Coxeter Presentation}

To present the group $\Sn$, the primitive operations we are going to use will be adjacent swaps. When dealing with
permutations on an $n + 1$-element set, there are $n$ adjacent transpositions, transposition number $k$ swapping
elements at indices $k$ and $k+1$. Thus, the generating set would be $\Fin[n]$. There are three relations that we're going
to specify for this presentation -- these are the equations that these generators should satisfy. We visualise them as
braid diagrams in~\Cref{fig:coxeter-braid}.

\begin{enumerate}
  \item[\labelcref{fig:coxeter-a}] Swapping the same two elements two times in a row should be the same as doing nothing
  \item[\labelcref{fig:coxeter-b}] When swapping two distinct pairs of elements, it should not matter in which order the
        swapping happens, that is, we can slide the wires freely.
  \item[\labelcref{fig:coxeter-c}] There are two equivalent ways of swapping the first and the last element in a
  three-element sequence.
\end{enumerate}

@@

Because of this counter-example, the relations have to be changed. In this section, we formally define a rewriting
system $(\List[\Fin[n]], \longcox)$, partially based on the Coxeter relations, and prove that it has the desired
properties of confluence and termination. We prove that the new relation defined by this system is equivalent,
in a technical sense, to the standard Coxeter relation $\cox*$.

First, we need to define a function $n \downf k$.

@@

The result of this function is the sequence \([k + n - 1, k + n - 2, k + n - 3, \ldots, n]\). The role of this function
is to produce a sequence of transpositions which moves element at index $k + n$ left by $k$ places, shifting all the
elements in between one place right (see~\cref{fig:downf}). Then, the directed relation $\longcox$ is defined with the
following generators, inlining the congruence closure in $\longcox$, allowing arbitrary reductions inside the list.

@@

Constructors $\longcancel$ and $\longswap$ correspond directly to the appropriate constructors of $\cox$ and can be
visualised in the same way as before. The remaining constructor $\longbraid$ uses the $\downf$ function to exchange the
order of a long sequence of transpositions and a single transposition afterwards. For example, for $n = 0$ and $k = 3$,
it allows for the reduction $[4, 3, 2, 1, 0, 4] \longcox [3, 4, 3, 2, 1, 0]$ (see~\cref{fig:longcox} -- note the
distinction between wires, where numbers represent the values, and transpositions, where numbers represent which wires
are crossing).

@@

      \def\nstrandsbr{6}
      \pic[local bounding box=my braid,braid/.cd,
        number of strands = \nstrandsbr,
        width = 0.4cm,
        height = 0.4cm,
        border height = 0.1cm,
        thick,
        name prefix=braid] at (4, 0)
      {braid={s_1, s_2, s_3, s_4, s_5, s_4}};
      % \draw[thick] % draws the top/bottom bars
      % ([xshift=-1ex]my braid.north west) --  ([xshift=1ex]my braid.north east)
      % ([xshift=-1ex]my braid.south west) --  ([xshift=1ex]my braid.south east);
      \foreach \n in {1,...,\nstrandsbl}{
          \pgfmathtruncatemacro{\nminusone}{\n - 1}
          \node at (braid-\n-s)[yshift = 2.8cm] {\nminusone};
        }
      % labels the bottom bar
      \node at (braid-1-e)[yshift = -2.8cm] {3};
      \node at (braid-2-e)[yshift = -2.8cm] {5};
      \node at (braid-3-e)[yshift = -2.8cm] {0};
      \node at (braid-4-e)[yshift = -2.8cm] {1};
      \node at (braid-5-e)[yshift = -2.8cm] {4};
      \node at (braid-6-e)[yshift = -2.8cm] {2};
    \end{tikzpicture}
    \caption{ $\longbraid$ with $n = 0$ and $k = 3$}
    \label{fig:longcox}
  \end{subfigure}
  \caption{Braiding diagrams for modified Coxeter relations.}
  \label{fig:coxeter-braid-mod}
\end{figure}

Note that the previous $\braid$ rule is a special case of $\longbraid$, with $k = 0$. Also, as before, the left-hand
sides of the relation are strictly larger than the right-hand sides, in terms of the lexicographic order on
$\List[\Fin[n]]$.

@@

For the relation $\longcox$, we define its transitive closure to be $\longcoxplus$ (\cref{def:longcox-plus}) and its
reflexive-transitive closure to be $\longcoxstar$ (\cref{def:longcox-star}).

Despite the increased complexity of the generators, the rewriting system $(\List[\Fin[n]],\longcox)$ has the properties
we desire. It satisfies local confluence, that is, the Church-Rosser (diamond) property (for example, using
$\longbraid$, we can now show the problematic critical pair \cref{fig:critical-pairs-non-converging} to converge), and it is
terminating, so, Newman's lemma, it produces a unique normal form. We follow the terminology
of~\cite*{huetConfluentReductionsAbstract1980a,krausCoherenceWellFoundednessTaming2020} to state our results formally.

@@

By~\Cref{prop:nf}~\ref{prop:nf-uniqueness}, we get a unique choice function $\normf : {\List[\Fin[n]]} \to
  {\List[\Fin[n]]}$ that produces a normal form for terms of $\List[\Fin[n]]$. We state and prove a couple of important
properties enjoyed by $\normf$.

@@

Finally, we define the type $\Sn$ as the set-quotient of $\List[\Fin[n]]$ by $\cox*$,

@@

Note that this relation $\cox*$ is not prop-valued, that is, $w \cox* v$ is not a proposition because reductions need
not be unique. So, the quotient $\Sn$ is not effective, that is, $\quotrel : w \cox* v \to q(w) \id q(v)$ is not an
equivalence. Using the $\normf$ function, we could instead define a new relation $\approx$ to equate those terms that
have the same normal form, $(w \approx v) \defeq (\normf(w) \id \normf(v))$. This relation is prop-valued, and we could
quotient $\List[\Fin[n]]$ by $\approx$, obtaining an equivalent definition for $\Sn$.

@@

\subsection{Lehmer Codes}~\label{subsec:lehmer}

@@

Formally, we define $\Lehmer[n]$ to be an $n+1$-element tuple, where the position $k \leq n$ stores an element of
$\Fin[k]$. Since the 0-th position is trivial, in practice it is ignored, and the sequence starts at
1~\cite{duboisTestsProofsCustom2018,vajnovszkiNewEulerMahonian2011}.

@@

From knowing just the inversion counts for all the elements, one can reconstruct the starting permutation. Also, observe
that $\invcount{i} < i$, thus fitting in the $i$-th place of a Lehmer code tuple. As an example, consider the following
tabulated presentation of a permutation of $\Fin[5]$.
\[
  \sigma =
  \begin{pmatrix}
    0      & 1      & 2      & 3      & 4      \\
    \el{2} & \el{1} & \el{4} & \el{0} & \el{3} \\
  \end{pmatrix}
\]

@@

The Lehmer code for the permutation $\sigma$ is then the 5-tuple
\[
  l = (\invcount{0}, \invcount{1}, \invcount{2}, \invcount{3}, \invcount{4}) = (0, 1, 2, 0, 2)
\]

To decode the permutation back from this Lehmer code, we perform an algorithm similar to \emph{insertion sort}. The
element of the Lehmer code being currently processed is highlighted in the left column of the table below. Starting from
a sorted list, the element at index $k$ has to be given $l[k]$ inversions. Because of the the invariant that all the
elements before newly processed one are smaller than it, the proper number of inversions is created by simply shifting
the element $l[k]$ places left.

@@

We can show that the function $\immersion_{n}$ gives an equivalence betweeen $\Lehmer[n]$ and $\im{\normf}$.

@@

Finally, it is time to complete our goal of characterising the permutation groups. Having produced a Lehmer code by
normalising words in $\Sn$, we need to run it to produce a concrete bijection of finite sets, and, given a bijection
between finite sets, we need to encode it as a Lehmer code. We will prove that these maps construct an equivalence
between the types $\Lehmer[n]$ and $\Aut[\Fin[\suc[n]]]$. The idea of this proof is borrowed from~\cite{Molzer-cubical}
~\footnote{Note that the indices for the type of permutations are off-by-one, because we chose $\Fin[n-1]$ to represent
  generators for permutations on $\Fin[n]$.}

\begin{definition}[${\FinExcept{n}} : {\Fin[n]} \to \UU$]
  For $n : \Nat$, the type family $\FinExcept{n}$ picks out all elements in $\Fin[n]$ except the one in the argument.
  \[
    \FinExcept{n}[i] \defeq \dsum{j : \Fin[n]}{i \neq j} \).
  \]
\end{definition}

Note that $\FinExcept{n}[i]$ for $i : \Fin[n]$ is a subtype of $\Fin[n]$ and is hence a set. We state and prove a few
auxiliary lemmas about how $\FinExcept{n}$ interacts with $\Fin$ -- these are obtained by counting arguments using the
fact that $\Fin[n]$ and $\FinExcept{n}$ both have decidable equality.

\begin{lemmarep}
  \label{prop:fin-finexcept}
  \leavevmode
  \begin{enumerate}
    % \item For any $k : \Fin[n]$, $\unit \sqcup \FinExcept{n}[k] \eqv \Fin[n]$. \label{prop:fin-finexcept-1}
    \item For any $k : \Fin[\suc[n]]$, $\FinExcept{\suc[n]}[k] \eqv \Fin[n]$. \label{prop:fin-finexcept-2}
    \item For any $n : \Nat$, \( \Aut[\Fin[\suc[n]]] \eqv \dsum{k : \Fin[\suc[n]]}{\FinExcept{\suc[n]}[n] \eqv
            \FinExcept{\suc[n]}[n - k]} \). \label{prop:fin-finexcept-3}
  \end{enumerate}
\end{lemmarep}

@@

Using these facts, we can now prove the main result of this section.

@@

\begin{proof}
  For $n = \zero$, note that $\Lehmer[\zero]$ is contractible, and so is $\Aut[\Fin[\suc[\zero]]]$. For $n = \suc[m]$,
  we compute a chain of equivalences.
  \begin{gather*}
    \arraycolsep=0.5em\def\arraystretch{1.5}
    \begin{array}{rl}
           & \Lehmer[\zero]          \\
      \eqv & \unit                   \\
      \eqv & \Aut[\Fin[\suc[\zero]]]
    \end{array}
    \qquad\qquad
    \begin{array}{rlr}
           & \Lehmer[\suc[m]]                                                                                    &                                                              \\
      \eqv & \Fin[\suc[\suc[m]]] \times \Lehmer[m]                                                               & \text{by definition}                                         \\
      \eqv & \Fin[\suc[\suc[m]]] \times \Aut[\Fin[\suc[m]]]                                                      & \text{induction hypothesis}                                  \\
      \eqv & \dsum*{k : \Fin[\suc[\suc[m]]]}{\Fin[\suc[m]] \eqv \Fin[\suc[m]]}                                   & \text{$\Sigma$ over a constant family}                       \\
      \eqv & \dsum*{k : \Fin[\suc[\suc[m]]]}{\FinExcept{\suc[\suc[m]]}[m] \eqv \Fin[\suc[m]]}                    & \text{\cref{prop:fin-finexcept}~\cref{prop:fin-finexcept-2}} \\
      \eqv & \dsum*{k : \Fin[\suc[\suc[m]]]}{\FinExcept{\suc[\suc[m]]}[m] \eqv \FinExcept{\suc[\suc[m]]}[m - k]} & \text{\cref{prop:fin-finexcept}~\cref{prop:fin-finexcept-2}} \\
      \eqv & \Aut[\Fin[\suc[\suc[m]]]]                                                                           & \text{\cref{prop:fin-finexcept}~\cref{prop:fin-finexcept-3}}
    \end{array}
  \end{gather*}
\end{proof}

By composing~\Cref{prop:lehmer-aut-equiv} and~\Cref{prop:sn-im-lehmer-equiv}, we obtain the final equivalence.

>>>>

We will show that it is possible to encode all $\PiLang$ combinators using adjacent transpositions
(in~\cref{sec:equivalence}). Group presentations are built by adding equations to a \emph{free group}.

@@

Before, the only way to decide the equality of two elements in a group was to evaluate and check them on the nose, but
in a group presentation, this is reduced to deciding whether one word -- a representative of the equivalence class of
the group's elements, can be reduced to another word, using the group's relations. However, these equations are not
directed, so it is not always possible to construct a well-behaved rewriting system. In general, the word problem for
groups is undecidable.

\paragraph{Coxeter Presentation} To present the group $\Sn$, the primitive operations we use will be adjacent swaps.
When dealing with permutations on an $n + 1$-element set, there are $n$ adjacent transpositions -- transposition number
$k$ swaps elements at indices $k$ and $k+1$. Thus, the generating set is $\Fin[n]$. There are three relations that we're
going to specify for this presentation -- we visualise them as braid diagrams in~\Cref{fig:coxeter-braid}.

\begin{enumerate}[leftmargin=*]
  \item[\labelcref{fig:coxeter-a}] Swapping the same two elements twice in a row should be the same as doing nothing.
  \item[\labelcref{fig:coxeter-b}] When swapping two distinct pairs of elements, the order in which swapping happens
        should not matter, that is, we can slide the wires freely.
  \item[\labelcref{fig:coxeter-c}] There are two equivalent ways of swapping the first and last elements in a sequence
        of three elements.
\end{enumerate}

@@

Because of this counter-example, the relations have to be changed. In this section, we formally define a rewriting
system $(\List[\Fin[n]], \longcox)$, partially based on the Coxeter relations, and prove that it has the desired
properties of confluence and termination. We prove that the new relation defined by this system is equivalent, in a
technical sense, to the standard Coxeter relation $\cox*$. First, we need to define a function $n \downf k$.

@@

\noindent The result of this function is the sequence \([k + n - 1, k + n - 2, k + n - 3, \ldots, n]\), which is a
sequence of transpositions that moves the element at index $k + n$ left by $k$ places, shifting all the elements in
between one place right (see~\cref{fig:downf}). Then, the directed relation $\longcox$ is defined with the following
generators (inlining the congruence closure in $\longcox$, allowing arbitrary reductions inside the list).

@@

\noindent Constructors $\longcancel$ and $\longswap$ correspond directly to the appropriate constructors of $\cox$ and
can be visualised in the same way as before. The remaining constructor $\longbraid$ uses the $\downf$ function to
exchange the order of a long sequence of transpositions and a single transposition afterwards. For example, for $n = 0$
and $k = 3$, it allows for the reduction $[4, 3, 2, 1, 0, 4] \longcox [3, 4, 3, 2, 1, 0]$ (see~\cref{fig:longcox} --
note the distinction between wires, where numbers represent the values, and transpositions, where numbers represent
which wires are crossing).

@@

      \def\nstrandsbr{6}
      \pic[local bounding box=my braid,braid/.cd,
        number of strands = \nstrandsbr,
        width = 0.4cm,
        height = 0.4cm,
        border height = 0.1cm,
        thick,
        name prefix=braid] at (4, 0)
      {braid={s_1, s_2, s_3, s_4, s_5, s_4}};
      % \draw[thick] % draws the top/bottom bars
      % ([xshift=-1ex]my braid.north west) --  ([xshift=1ex]my braid.north east)
      % ([xshift=-1ex]my braid.south west) --  ([xshift=1ex]my braid.south east);
      \foreach \n in {1,...,\nstrandsbl}{
          \pgfmathtruncatemacro{\nminusone}{\n - 1}
          \node at (braid-\n-s)[yshift = 2.8cm] {\nminusone};
        }
      % labels the bottom bar
      \node at (braid-1-e)[yshift = -2.8cm] {3};
      \node at (braid-2-e)[yshift = -2.8cm] {5};
      \node at (braid-3-e)[yshift = -2.8cm] {0};
      \node at (braid-4-e)[yshift = -2.8cm] {1};
      \node at (braid-5-e)[yshift = -2.8cm] {4};
      \node at (braid-6-e)[yshift = -2.8cm] {2};
    \end{tikzpicture}
    \caption{ $\longbraid$ with $n = 0$ and $k = 3$}
    \label{fig:longcox}
  \end{subfigure}
  \caption{Braiding diagrams for modified Coxeter relations.}
  \label{fig:coxeter-braid-mod}
\end{figure}
% in terms of the lexicographic order on $\List[\Fin[n]]$.

@@

Note that the previous $\braid$ rule is a special case of $\longbraid$, with $k = 0$. As before, the left-hand sides of
the relation are (lexicographically) strictly larger than the right-hand sides. We define the transitive closure of
$\longcox$ to be $\longcoxplus$ (\cref{def:longcox-plus}) and its reflexive-transitive closure to be $\longcoxstar$
(\cref{def:longcox-star}).

Despite the increased complexity of the generators, the rewriting system $(\List[\Fin[n]],\longcox)$ has the properties
we desire. It satisfies local confluence, that is, the Church-Rosser (diamond) property -- for example, using
$\longbraid$, we can now show the problematic critical pair \cref{fig:critical-pairs-non-converging} converges, and, it
is terminating, so by Newman's lemma, it produces a unique normal form. We follow the terminology
of~\citet*{huetConfluentReductionsAbstract1980a,krausCoherenceWellFoundednessTaming2020} to state our results formally.

@@

By~\Cref{prop:nf}~\ref{prop:nf-uniqueness}, we get a unique choice function $\normf : {\List[\Fin[n]]} \to
  {\List[\Fin[n]]}$ that produces a normal form for terms of $\List[\Fin[n]]$. We state two important properties enjoyed
by $\normf$.

@@

\noindent Finally, we define the type $\Sn$ as the set-quotient of $\List[\Fin[n]]$ by $\cox*$,

@@

\noindent Note that reductions need not be unique, hence $w \cox* v$ is not necessarily a proposition. So, the quotient
$\Sn$ is not effective, that is, $\quotrel : w \cox* v \to q(w) \id q(v)$ is not an equivalence. Using the $\normf$
function, we could instead define a new relation $\approx$ to equate those terms that have the same normal form, $(w
  \approx v) \defeq (\normf(w) \id \normf(v))$. This relation is prop-valued, and we could quotient $\List[\Fin[n]]$ by
$\approx$, obtaining an equivalent definition for $\Sn$.

@@

\subsection{Lehmer Codes}
\label{subsec:lehmer}

@@

Formally, we define $\Lehmer[n]$ to be an $n+1$-element tuple, where the position $k \leq n$ stores an element of
$\Fin[k]$. Since the 0-th position is trivial, in practice it is
ignored~\cite{duboisTestsProofsCustom2018,vajnovszkiNewEulerMahonian2011}.

@@

Just knowing the inversion counts for all the elements, one can reconstruct the starting permutation. Also, observe that
$\invcount{i} < i$, thus fitting in the $i$-th place of a Lehmer code tuple. As an example, consider the following
tabulated presentation of a permutation of $\Fin[5]$.
\[
  \sigma =
  \begin{pmatrix}
    0      & 1      & 2      & 3      & 4      \\
    \el{2} & \el{1} & \el{4} & \el{0} & \el{3} \\
  \end{pmatrix}
\]

@@

\noindent The Lehmer code for the permutation $\sigma$ is then the 5-tuple
\[
  l = (\invcount{0}, \invcount{1}, \invcount{2}, \invcount{3}, \invcount{4}) = (0, 1, 2, 0, 2)
\]

\noindent To decode the permutation back from this Lehmer code, we perform an algorithm similar to \emph{insertion
  sort}. The element of the Lehmer code being currently processed is highlighted in the left column of the table below.
Starting from a sorted list, the element at index $k$ has to be given $l[k]$ inversions. Because of the the invariant
that all the elements before newly processed one are smaller than it, the proper number of inversions is created by
simply shifting the element $l[k]$ places left.

@@

\noindent We can show that the function $\immersion_{n}$ gives an equivalence betweeen $\Lehmer[n]$ and $\im{\normf}$.

@@

Finally, it is time to complete our goal of characterising the permutation groups. Having produced a Lehmer code by
normalising words in $\Sn$, we need to run it to produce a concrete bijection of finite sets, and, given a bijection
between finite sets, we need to encode it as a Lehmer code. We will prove that these maps construct an equivalence
between the types $\Lehmer[n]$ and $\Aut[\Fin[\suc[n]]]$. The idea for this proof is borrowed
from~\cite{Molzer-cubical}. ~\footnote{Note that the indices for the type of permutations are off-by-one, because we
  chose $\Fin[n]$ to represent generators for permutations on $\Fin[\suc[n]]$.}

\begin{definition}[${\FinExcept{n}} : {\Fin[n]} \to \UU$]
  For $n : \Nat$, the type family $\FinExcept{n}$ picks out all elements in $\Fin[n]$ except the one in the argument.
  \[
    \FinExcept{n}[i] \defeq \dsum*{j : \Fin[n]}{i \neq j}
  \]
\end{definition}

\noindent Note that $\FinExcept{n}[i]$ for $i : \Fin[n]$ is a subtype of $\Fin[n]$ and is hence a set. We state and
prove a few auxiliary lemmas about how $\FinExcept{n}$ interacts with $\Fin$ -- these are obtained by counting arguments
using the fact that $\Fin[n]$ and $\FinExcept{n}$ both have decidable equality.

\begin{lemmarep}
  \label{prop:fin-finexcept}
  \leavevmode
  \begin{enumerate}
    % \item For any $k : \Fin[n]$, $\unit \sqcup \FinExcept{n}[k] \eqv \Fin[n]$. \label{prop:fin-finexcept-1}
    \item For any $k : \Fin[\suc[n]]$, we have $\FinExcept{\suc[n]}[k] \eqv \Fin[n]$. \label{prop:fin-finexcept-2}
    \item For any $n : \Nat$, we have \( \Aut[\Fin[\suc[n]]] \eqv \dsum*{k : \Fin[\suc[n]]}{\FinExcept{\suc[n]}[n] \eqv
            \FinExcept{\suc[n]}[n - k]} \). \label{prop:fin-finexcept-3}
  \end{enumerate}
\end{lemmarep}

@@

\noindent Using these facts, we can now prove the main result of this section.

@@

\begin{proof}
  For $n = \zero$, note that $\Lehmer[\zero]$ is contractible, and so is $\Aut[\Fin[\suc[\zero]]]$. For $n = \suc[m]$,
  we compute a chain of equivalences.
  \begin{gather*}
    \arraycolsep=0.5em\def\arraystretch{1.5}
    \begin{array}{rl}
           & \Lehmer[\zero]          \\
      \eqv & \unit                   \\
      \eqv & \Aut[\Fin[\suc[\zero]]]
    \end{array}
    \qquad\qquad
    \begin{array}{rlr}
           & \Lehmer[\suc[m]]                                                                                    &                                                                 \\
      \eqv & \Fin[\suc[\suc[m]]] \times \Lehmer[m]                                                               & \text{by definition}                                            \\
      \eqv & \Fin[\suc[\suc[m]]] \times \Aut[\Fin[\suc[m]]]                                                      & \text{induction hypothesis}                                     \\
      \eqv & \dsum*{k : \Fin[\suc[\suc[m]]]}{\Fin[\suc[m]] \eqv \Fin[\suc[m]]}                                   & \text{$\Sigma$ over a constant family}                          \\
      \eqv & \dsum*{k : \Fin[\suc[\suc[m]]]}{\FinExcept{\suc[\suc[m]]}[m] \eqv \Fin[\suc[m]]}                    & \text{by~\cref{prop:fin-finexcept}~\cref{prop:fin-finexcept-2}} \\
      \eqv & \dsum*{k : \Fin[\suc[\suc[m]]]}{\FinExcept{\suc[\suc[m]]}[m] \eqv \FinExcept{\suc[\suc[m]]}[m - k]} & \text{by~\cref{prop:fin-finexcept}~\cref{prop:fin-finexcept-2}} \\
      \eqv & \Aut[\Fin[\suc[\suc[m]]]]                                                                           & \text{by~\cref{prop:fin-finexcept}~\cref{prop:fin-finexcept-3}}
    \end{array}
  \end{gather*}
\end{proof}

\noindent By composing~\Cref{prop:lehmer-aut-equiv} and~\Cref{prop:sn-im-lehmer-equiv}, we obtain the final equivalence.

### 2021-07-09 Pi+/paper/ufin.tex
Futher work

We have shown that loops in $\UFin$ exactly encode the automorphism group $\Aut[\Fin[n]]$ for every~$n$. This is a
general technique called \emph{delooping}, where a group can be identified with a 1-object groupoid, internally in
HoTT. This technique also allows defining higher groups in HoTT~\cite{buchholtzHigherGroupsHomotopy2018}. The loopspace
of a pointed type automatically has the structure of a group, with $\refl_{\pt}$ for the neutral element, path
composition for the group multiplication, and path inverse for the group inverse operation. The group axioms are then
given by the higher paths corresponding to groupoid laws.

@@

\begin{theorem}
  $\UFin$ has a symmetric rig structure, with the additive and multiplicative symmetric monoidal structures given by
  $(F_0, \sqcup)$ and $(F_1, \times)$, with corresponding natural isomorphisms $\lambda_{X}$, $\rho_{X}$,
  $\alpha_{X,Y,Z}$, and the braiding isomorphism $\mathcal{B}_{X,Y}$ upto 1-paths in $\UFin$. These isomorphisms satisfy
  the Mac Lane coherence conditions for symmetric monoidal categories, that is, the triangle, pentagon, and hexagon
  identities, and the symmetry of the braiding, upto 2-paths in $\UFin$. \todo{Add rig}
\end{theorem}

>>>>

We have shown that loops in $\UFin$ exactly encode the automorphism group $\Aut[\Fin[n]]$ for every~$n$. This is a
general technique called \emph{delooping}, where a group can be identified with a 1-object groupoid, internally in HoTT.
This technique also allows defining higher groups~\cite{buchholtzHigherGroupsHomotopy2018}. The loopspace of a pointed
type automatically has the structure of a group, with $\refl_{\pt}$ for the neutral element, path composition for the
group multiplication, and path inverse for the group inverse operation. The group axioms are given by the higher paths
corresponding to groupoid laws.

@@

\begin{theorem}
  $\UFin$ has two symmetric monoidal structures, the additive and multiplicative ones, given by $(F_0, \sqcup)$ and
  $(F_1, \times)$, with corresponding natural isomorphisms $\lambda_{X}$, $\rho_{X}$, $\alpha_{X,Y,Z}$, and the braiding
  isomorphism $\mathcal{B}_{X,Y}$ upto 1-paths in $\UFin$. These isomorphisms satisfy the Mac Lane coherence conditions
  for symmetric monoidal categories~\cite{maclaneNaturalAssociativityCommutativity1963}, that is, the triangle,
  pentagon, and hexagon identities, and the symmetry of the braiding, upto 2-paths in $\UFin$. The multiplicative
  structure distributes over the additive structure and satisfies the Laplaza coherence conditions for rig
  categories~\cite{laplaza72}.
\end{theorem}
